---
title: "P8105_hw3_cz2750"
output: github_document
date: "2023-10-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE) 
library(p8105.datasets)
library(tidyverse)
library(knitr)
```

# Q1

```{r}
data("instacart")
```


## 1. Dataset Description:
1. The dataset `instacart` has `r nrow(instacart)` observations and `r ncol(instacart)` variables.
2. Variables in the dataset include `r colnames(instacart)`.
3. Key variables:
  1). order_id: order identifier
  2). add_to_cart_order: order in which each product was added to cart
  3). reordered: 1 if this prodcut has been ordered by this user in the past, 0 otherwise
  4). order_dow: the day of the week on which the order was placed
  5). product_name: name of the product
  
## 2. Number of Aisles and which aisles are the most items ordered from
```{r}
number <- n_distinct(instacart$aisle)
aisle_counts <- instacart %>%
  group_by(aisle) %>%
  summarise(count = n()) %>%
  arrange(-count) %>%
  slice(1) %>%
  pull(aisle) 
```
1). There are 'r number' unique aisles.
2). The most items are ordered from the 'aisle_counts'.

## 3. Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
aisle_counts_filtered <- instacart %>%
  group_by(aisle) %>%
  summarise(count = n()) %>%
  filter(count > 10000) %>%
  arrange(-count)

ggplot(aisle_counts_filtered, aes(x = reorder(aisle, count), y = count)) +
  geom_bar(stat="identity") +
  coord_flip() + 
  labs(title = "Number of Items Ordered in Each Aisle",
       x = "Aisle",
       y = "Total Items Ordered") +
  theme_minimal()
```

## 4. Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
popitems <- instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle, product_name) %>%
  summarise(count = n(), .groups = "drop") %>%
  arrange(aisle, -count) %>%
  group_by(aisle) %>%
  slice_head(n = 3)

knitr::kable(popitems,caption = "Top 3 Items in Selected Aisles", align = c('l', 'l', 'r'))
```
## 5. Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
prohour <- instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarise(mean_hour = mean(order_hour_of_day, na.rm = TRUE), .groups = "drop") %>%
  spread(key = order_dow, value = mean_hour) %>%
  select(product_name, `0`, `1`, `2`, `3`, `4`, `5`, `6`) 

knitr::kable(prohour, caption = "Mean Hour of Order by Product and Day of the Week", digits = 2)
```
# Q2

```{r}
data("brfss_smart2010")
```


## 1. Data Cleaning
```{r}
colnames(brfss_smart2010)
brfss_smart2010 <- brfss_smart2010 %>%
  rename(
    class_id = ClassId,
    data_source = DataSource,
    geo_location = GeoLocation,
    location_abbr = Locationabbr,
    location_desc = Locationdesc,
    location_id = LocationID,
    question_id = QuestionID,
    resp_id = RESPID,
    topic_id = TopicId,
  ) %>%
    janitor::clean_names() %>%
  filter(topic == "Overall Health",
         response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>%
  mutate(response = factor(response, 
                           levels = c("Poor", "Fair", "Good", "Very good", "Excellent"),
                           ordered = TRUE))
```
## 2. States Observed at 7 or More Locations in 2002 and 2010
```{r}
states <- brfss_smart2010 %>%
  filter(year %in% c(2002, 2010)) %>%
  group_by(year, location_abbr) %>%
  summarise(num_entries = n_distinct(location_desc)) %>%
  filter(num_entries >= 7)

states_2002 <- states %>%
  filter(year == 2002)  %>%
  pull(location_abbr)

states_2010 <- states %>%
  filter(year == 2010) %>%
  pull(location_abbr)
```
1). In 2002, states observed at 7 or more locations were `r paste(states_2002, collapse=", ")`.
2). In 2010, states observed at 7 or more locations were `r paste(states_2010, collapse=", ")`.

## 3. Construct a dataset and create a Spaghetti Plot for Excellent Responses
```{r}
excellent_responses <- brfss_smart2010 %>%
  filter(response == "Excellent") %>%
  group_by(year, location_abbr) %>%
  summarise(avg_data_value = round(mean(data_value, na.rm = TRUE),2))

ggplot(excellent_responses, aes(x = year, y = avg_data_value, group = location_abbr)) +
  geom_line(aes(color = location_abbr), alpha = 0.7) +
  labs(title = "Average value trends for each state",
       y = "Average Data Value") +
  theme_minimal() 
```

## 4. Two-Panel Plot for Distribution of Data Values in NY State (2006 and 2010)
```{r}
ny_data <- brfss_smart2010 %>%
  filter(year %in% c(2006, 2010) & location_abbr == "NY")

ggplot(ny_data, aes(x = response, y = data_value)) +
  geom_boxplot() +
  facet_wrap(~ year, ncol = 2) +
  labs(title = "Distribution of Data Value for responses in NY (2006 & 2010)",
       y = "Data Value",
       x = "Response") +
  theme_minimal()
```

